{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-01 thesis/tm.tex
more tm revisions and 10hr resulce

Intel's official TSX documentation summarizes its interface and behaviour in prose \cite{intel-tsx-overview}.
Recent work has used proof assistants to formalize some of the execution semantics
of x86 in general \cite{x86-semantics}
and of transactions in particular (at both hardware- and language-level) \cite{relaxed-transactions-pldi}.
However, state-of-the-art advances in model checking algorithms
still state their theorems and proofs in prose \cite{bpor,optimal-dpor,tsopso,satcheck,mcr,quicksand,cpp-model-checking},
so this section's proofs will regrettably do likewise,
leaving the rigor of mechanization to future work.
The reader may at least rest assured that the proofs herein
rely on transactional semantics that have themselves been formally verified.

@@

To address this,
I extended Landslide
with {\em retry set reduction} (commit 86657c7),
an experimental feature named after prior work's analogous {\em sleep sets}
(\cite{partial-order-methods,dpor,optimal-dpor}, \sect{\ref{sec:landslide-sleepsets}}).
Whenever DPOR tags a new branch for exploration (\sect{\ref{sec:landslide-explore}}),
if either or both sides of a memory conflict were part of a transaction,
it records a {\em retry set},
i.e., the pair of {\tt xbegin} results executed by the conflicting threads,
to accompany that branch.
Like the {\tt xabort\_codes} lists, the state space estimators (\sect{\ref{sec:landslide-estimate}})
check the number of retry sets when counting the number of marked children.
% ah-- not compatible after all!
%in addition to tracking the set of abort codes possible at each preemption point.
Later, when traversing the new branch,
Landslide remembers those threads' expected {\tt xbegin} results
and refuses to test any others
(unless DPOR separately found them, too, to conflict),
thereby skipping over (i.e., pruning) reorderings of any other {\tt xbegin} results that would be independent.
%The {\tt landslide} repository's {\tt retry-sets} branch implements this feature,
Upcoming in the evaluation, Figure~\ref{fig:retry-reduction} will visualize the reduction achieved in two test cases.

@@

			& 2,4 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq} \\

@@

			& 3,2 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq} \\

@@

			& 3,1 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	&\ETAdag{97y 100d}&\ETAdag{1698185036} \\

@@

			& 2,2 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq} \\
			& 3,1 & \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq}	& \ETAdag{qqq} \\

>>>>

Intel's official TSX documentation summarizes its interface and behaviour in prose \cite{intel-tsx-overview}.
Recent work has used proof assistants to formalize some of the execution semantics
of x86 in general \cite{x86-semantics}
and of transactions in particular (at both hardware- and language-level) \cite{relaxed-transactions-pldi}.
However, state-of-the-art advances in model checking algorithms
still state their theorems and proofs in prose \cite{bpor,optimal-dpor,tsopso,satcheck,mcr,quicksand,rcmc},
so this section's proofs will regrettably do likewise,
leaving the rigor of mechanization to future work.
The reader may at least rest assured that the proofs herein
rely on transactional semantics that have themselves been formally verified.

@@

To address this,
I extended Landslide
with {\em retry set reduction} (commit 86657c7),
an experimental feature named after prior work's analogous {\em sleep sets}
(\cite{partial-order-methods,dpor,optimal-dpor}, \sect{\ref{sec:landslide-sleepsets}}).
Whenever DPOR tags a new branch for exploration (\sect{\ref{sec:landslide-explore}}),
if either or both sides of a memory conflict were part of a transaction,
it records a {\em retry set},
i.e., the pair of {\tt xbegin} results executed by the conflicting threads,
to accompany that branch.
Like the {\tt xabort\_codes} lists, the state space estimators (\sect{\ref{sec:landslide-estimate}})
check the number of retry sets when counting the number of marked children.
% ah-- not compatible after all!
%in addition to tracking the set of abort codes possible at each preemption point.
Later, when traversing the new branch,
Landslide remembers those threads' expected {\tt xbegin} results
and refuses to test any others
(unless DPOR separately found them, too, to conflict),
thereby skipping over (i.e., pruning) reorderings of any other {\tt xbegin} results that would be independent.
Like sleep sets, it also considers the preempted thread ``retry-set blocked''
(like sleep-set blocking \cite{optimal-dpor}),
and refuses to run it until the conflicting thread runs its transaction first,
unless such would result in deadlock%
\footnote{
	Before the search ordering update to Landslide's normal sleep sets implementation
	(\sect{\ref{sec:landslide-sleepsets}}),
	I observed false-positive retry-set-blocked deadlocks fairly often;
	after the update,
	it took until 981773 interleavings (>9 hours) into {\tt htm2}(3,2) to find one
	and confirm the need to still explicitly avoid them by abandoning the retry set in such cases.
	% TODO put a commit # here
}.
%The {\tt landslide} repository's {\tt retry-sets} branch implements this feature,
Upcoming in the evaluation, Figure~\ref{fig:retry-reduction} will visualize the reduction achieved in two test cases.

@@

			& 2,4 & \ETAdag{29d 8h}	&\ETAdag{2710056}&\cpu{4771.17}	& \ints{123140}	&\ETAdag{5d 1h}	&\ETAdag{1792330} \\
			% [JOB 4] progress: 683895/2710056 brs (25.235453%), ETA 28d 21h 54m 32s (elapsed 9h 59m 54s)
			% [JOB 4] progress: 714585/1792330 brs (39.869037%), ETA 4d 14h 46m 38s (elapsed 9h 59m 54s)

@@

			& 3,2 &\ETAdag{1y 299d}	&\ETAdag{4510472}& \ETAdag{qqq}	& \ETAdag{qqq}	&\ETAdag{10d 16h}&\ETAdag{2322150}\\
			% [JOB 4] progress: 759740/4510472 brs (16.843911%), ETA 1y 298d 22h 14m 13s (elapsed 9h 59m 55s)
			% [JOB 4] progress: 646606/2322150 brs (27.845142%), ETA 10d 5h 50m 15s (elapsed 9h 59m 55s)

@@

			& 3,1 & \ETAdag{5y 292d}&\ETAdag{2873664642}&\ETAdag{qqq}& \ETAdag{qqq}	&\ETAdag{97y 100d}&\ETAdag{1698185036} \\
			% [JOB 18] progress: 447687/2873664642 brs (0.015579%), ETA 5y 291d 11h 13m 30s (elapsed 9h 59m 41s)

@@

			& 2,2 & \ETAdag{1y 176d}&\ETAdag{75516602}&\ETAdag{qqq}	& \ETAdag{qqq}	&\ETAdag{13d 11h}&\ETAdag{565334} \\
			% [JOB 22] progress: 87777/75516602 brs (0.116235%), ETA 1y 176d 16h 38m 50s (elapsed 3h 1m 48s)
			% [JOB 22] progress: 18380/565334 brs (3.251171%), ETA 13d 1h 5m 3s (elapsed 38m 33s)
			& 3,1 &\ETAdag{6y 133d}&\ETAdag{1847957714}&\ETAdag{qqq}& \ETAdag{qqq}	&\ETAdag{4y 156d}&\ETAdag{888242178} \\
			% [JOB 22] progress: 284530/1847957714 brs (0.015397%), ETA 6y 132d 6h 20m 26s (elapsed 9h 49m 5s)
			% [JOB 22] progress: 274532/888242178 brs (0.030907%), ETA 4y 155d 23h 50m 51s (elapsed 9h 47m 5s)
